---
title: "Average Temperature and Precipitation on main cities"
output: html_document
date: '2023-09-24'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup 

None

## Libraries

```{r echo=FALSE}
if (!"conflicted" %in% installed.packages()) install.packages("conflicted")
library(conflicted)

if (!"styler" %in% installed.packages()) install.packages("styler")
library(styler)

library("ecmwfr")
library(sf)
library(ncdf4)

if (!"tidyverse" %in% installed.packages()) install.packages("tidyverse")
library(tidyverse)
library(purrr)

if (!"fs" %in% installed.packages()) install.packages("fs")
library(fs)

library(lubridate, warn.conflicts = FALSE)
library(tidync)
library(stars)

if (!"maps" %in% installed.packages()) install.packages("maps")
library(maps)

if (!"ggmap" %in% installed.packages()) install.packages("ggmap")
library(ggmap)

if (!"mapview" %in% installed.packages()) install.packages("mapview")
library(mapview)

if (!"spData" %in% installed.packages()) install.packages("spData")
library(spData)

library(ggplot2)
library(hrbrthemes)

if (!"plotly" %in% installed.packages()) install.packages("plotly")
library(plotly)
if (!"gapminder" %in% installed.packages()) install.packages("gapminder")
library(gapminder)

if (!"ggtext" %in% installed.packages()) install.packages("ggtext")
library(ggtext)


if (!"svglite" %in% installed.packages()) install.packages("svglite")
library(svglite)

conflict_prefer("map", "purrr")
conflict_prefer("filter", "dplyr")

```

## Getting Data

We have generated the data from the CDS web site and store the results on the file -> See in code below

## Loading data

Reading data from ncdf file

```{r echo=FALSE}
# Reading Spanish cities data // With population > 30000. To get at least all the capitals of Spain
cities_set_df <- read_csv("./datasets/worldcities.csv") |>
  filter(iso2 == "ES", population > 30000, !is.na(capital) )

cities_set_sf <- cities_set_df %>%
  st_as_sf(coords = c("lng", "lat"), remove = FALSE)

st_crs(cities_set_sf) <- 4326

# OLD Climate Data Hadley Centre observations datasets

climatologies_world_HadCRUT_5_monthly_nc <- tidync("./datasets/HadCRUT.5.0.1.0.analysis.anomalies.ensemble_mean.nc")
climatologies_world_HadCRUT_baseline_1961_1990_nc <- tidync("./datasets/absolute_v5.nc") 
climatologies_world_HadCRUT_baseline_1961_1990_dat <- climatologies_world_HadCRUT_baseline_1961_1990_nc |>
  hyper_tibble() |>
  mutate(month = ceiling(time)) |>
  mutate(longitude = lon, latitude = lat)

print(nc_open("./datasets/HadCRUT.5.0.1.0.analysis.anomalies.ensemble_mean.nc"))
# Climate Data CDS

# # Equivalent CDS API Request
# import cdsapi
# 
# c = cdsapi.Client()
# 
# c.retrieve(
#     'reanalysis-era5-single-levels-monthly-means',
#     {
#         'product_type': 'monthly_averaged_reanalysis',
#         'variable': [
#             '2m_temperature', 'total_precipitation',
#         ],
#         'year': [
#             '1959', '1960', '1961',
#             '1962', '1963', '1964',
#             '1965', '1966', '1967',
#             '1968', '1969', '1970',
#             '1971', '1972', '1973',
#             '1974', '1975', '1976',
#             '1977', '1978', '1979',
#             '1980', '1981', '1982',
#             '1983', '1984', '1985',
#             '1986', '1987', '1988',
#             '1989', '1990', '1991',
#             '1992', '1993', '1994',
#             '1995', '1996', '1997',
#             '1998', '1999', '2000',
#             '2001', '2002', '2003',
#             '2004', '2005', '2006',
#             '2007', '2008', '2009',
#             '2010', '2011', '2012',
#             '2013', '2014', '2015',
#             '2016', '2017', '2018',
#             '2019', '2020', '2021',
#         ],
#         'month': [
#             '01', '02', '03',
#             '04', '05', '06',
#             '07', '08', '09',
#             '10', '11', '12',
#         ],
#         'time': '00:00',
#         'format': 'netcdf',
#     },
#     'download.nc')

climatologies_world_file <- "./datasets/adaptor.mars.internal-1664035655.475343-28557-17-dbf6e11e-c65f-421f-80ee-2dc637adc596.nc"

# Data pre 2022
climatologies_world_nc <- tidync(climatologies_world_file)

# Data 2022

# Equivalent CDS API Request
# import cdsapi
# 
# c = cdsapi.Client()
# 
# c.retrieve(
#     'reanalysis-era5-single-levels-monthly-means',
#     {
#         'product_type': 'monthly_averaged_reanalysis',
#         'variable': [
#             '2m_temperature', 'total_precipitation',
#         ],
#         'year': '2022',
#         'month': [
#             '01', '02', '03',
#             '04', '05', '06',
#             '07', '08',
#         ],
#         'time': '00:00',
#         'format': 'netcdf',
#     },
#     'download.nc')

climatologies_world_2022_nc <- tidync("./datasets/adaptor.mars.internal-1664443401.824954-13047-14-23dfdef7-7e28-4281-98f5-2c23ddcc11ce.nc")

##### DATA - E-OBS ####
climatologies_europe_eobs_1950_2022_nc <- tidync("./datasets/tg_ens_mean_0.1deg_reg_v25.0e.nc")

#lonlat_df <- climatologies_world_nc %>% hyper_filter(time = index == 1, latitude = latitude > 4 & latitude < 6, longitude = longitude > 4 & longitude < 6) %>% hyper_tibble()

# Build sf for the lon/lat values
lonlat_sf <- climatologies_world_nc %>% 
  hyper_filter(time = index == 1) %>% 
  hyper_tibble() %>% 
  select(longitude, latitude)  %>% 
  mutate(longitude = ifelse(longitude > 180, -360 + longitude, longitude))  %>% 
  st_as_sf(coords = c("longitude", "latitude"),remove = FALSE)
  
st_crs(lonlat_sf) <- 4326

# Find out closest coordinates indexes
nearest_coords <- st_nearest_feature(cities_set_sf, lonlat_sf)

# Get the values and reverting the longitudes to the original format. Because we will use them to filter the # whole original file
lon_lat_values <- lonlat_sf  |>  
  slice(nearest_coords) |>
  select(longitude,latitude) |> 
  mutate(longitude = ifelse(longitude < 0, +360 + longitude, longitude))  %>% 
  (\(x) list( lon = pull(x,longitude), lat = pull(x,latitude)))()
  
# Add closest coords in original format to our set of cities

cities_set_ex_df <- add_column(cities_set_sf, lon_orig = lon_lat_values$lon, lat_orig = lon_lat_values$lat)


get_time_slices_for_cities <- function(cities_set_ex_df, climatologies_world_nc) {
#' Get time slices from the gridded climatologies netCDF file for the cties in cities_set
#' 
#' Return an sf with simple points for all the coordinates, and some merged info from the cities set  
#'

  climatology_cities_dat <- cities_set_ex_df$city |> 
    map(function(city_name) {
          lon = cities_set_ex_df |> filter(city == city_name) |> pull(lon_orig) 
          lat = cities_set_ex_df |> filter(city == city_name) |> pull(lat_orig) 
          climatologies_world_nc |> 
            hyper_filter( longitude =  longitude == lon, latitude = latitude == lat) |>
            hyper_tibble() |>
            add_column(city = city_name)
        }) |>
    bind_rows() |>
    mutate(longitude = ifelse(longitude > 180, -360 + longitude, longitude)) |>
    mutate(t2m_c = t2m - 273.15) |> # convert Kelvin to Celsius
    mutate(date_time = as_datetime(c(time * 60 * 60), origin="1900-01-01")) |>
    mutate(year = year(date_time)) |>
    mutate(month = month(date_time)) |>
    mutate(month_name = month(date_time, label = TRUE)) |>
    mutate(decade = as.character(as.integer((year %/% 10) * 10))) |> # Add decade class
    st_as_sf(coords = c("longitude", "latitude"),remove = FALSE)
  
  st_crs(climatology_cities_dat) <- 4326
  climatology_cities_dat
}


climatology_cities_pre_2022_dat <- get_time_slices_for_cities(cities_set_ex_df, climatologies_world_nc)

climatology_cities_2022_dat <- get_time_slices_for_cities(cities_set_ex_df, climatologies_world_2022_nc)


#######################################################################
#
# Build Preindustrial Data
#
#######################################################################

# Build sf for the lon/lat values
lonlat_pre_industrial_sf <- climatologies_world_HadCRUT_5_monthly_nc |>
  hyper_filter(time = index == 1) |>
  hyper_tibble() |> 
  select(longitude, latitude) |>
  st_as_sf(coords = c("longitude", "latitude"), remove = FALSE)
  
st_crs(lonlat_pre_industrial_sf) <- 4326

# Find out closest coordinates indexes
nearest_coords_pre_industrial <- st_nearest_feature(cities_set_sf, lonlat_pre_industrial_sf)

# Get the values and reverting the longitudes to the original format. Because we will use them to filter the # whole original file
lon_lat_pre_industrial_values <- lonlat_pre_industrial_sf  |>  
  slice(nearest_coords_pre_industrial) |>
  select(longitude,latitude) |> 
  (\(x) list( lon = pull(x,longitude), lat = pull(x,latitude)))()
  
# Add closest coords in original format to our set of cities

cities_set_ex_df <- add_column(cities_set_ex_df, lon_orig_pre_industrial = lon_lat_pre_industrial_values$lon, lat_orig_pre_industrial = lon_lat_pre_industrial_values$lat)


get_time_slices_HadCRUT_5_for_cities <- function(cities_set_ex_df, climatologies_world_HadCRUT_5_nc) {
#' Get time slices from the gridded climatologies netCDF file for the cties in cities_set
#' 
#' Return an sf with simple points for all the coordinates, and some merged info from the cities set  
#'

  climatology_cities_dat <- cities_set_ex_df$city |> 
    map(function(city_name) {
          lon = cities_set_ex_df |> filter(city == city_name) |> pull(lon_orig_pre_industrial) 
          lat = cities_set_ex_df |> filter(city == city_name) |> pull(lat_orig_pre_industrial) 
          climatologies_world_HadCRUT_5_nc |> 
            hyper_filter( longitude =  longitude == lon, latitude = latitude == lat) |>
            hyper_tibble() |>
            add_column(city = city_name)
        }) |>
    bind_rows() |>
    mutate(t2m_c = tas_mean) |> # Difference in kelvin = difference in Celsius
    mutate(date_time = as_datetime(c(time * 60 * 60 * 24), origin="1850-01-01")) |>
    mutate(year = year(date_time)) |>
    mutate(month = month(date_time)) |>
    mutate(month_name = month(date_time, label = TRUE)) |>
    mutate(decade = as.character(as.integer((year %/% 10) * 10))) |> # Add decade class
    st_as_sf(coords = c("longitude", "latitude"),remove = FALSE)

  st_crs(climatology_cities_dat) <- 4326
  climatology_cities_dat
}

# Convert the anomalies to absolute temperature values
climatology_cities_pre_industrial_dat <- 
  get_time_slices_HadCRUT_5_for_cities(cities_set_ex_df, climatologies_world_HadCRUT_5_monthly_nc) |>
  left_join(climatologies_world_HadCRUT_baseline_1961_1990_dat,  by = c("longitude", "latitude", "month")) |>
  mutate(t2m_c = t2m_c + tem)


pre_industrial_baseline_dat <- climatology_cities_pre_industrial_dat |>
  filter(year < 1900) |>
  group_by(city, month) |>
  summarise(temperature_climatology_mean = mean(t2m_c)) |>
  as.tibble() |>
  select(-geometry)

climatology_cities_pre_industrial_baseline_dat <- climatology_cities_pre_industrial_dat |>
  left_join(pre_industrial_baseline_dat, by = c("city", "month"))



#######################################################################
#
# Build Climatology Europe From E-OBS - 1950 -2022
#
#######################################################################

# Build sf for the lon/lat values
climatologies_europe_eobs_1950_2022_lon_lat_sf <- climatologies_europe_eobs_1950_2022_nc |>
  hyper_filter(time = index == 1) |>
  hyper_tibble() |> 
  select(longitude, latitude) |>
  st_as_sf(coords = c("longitude", "latitude"), remove = FALSE)
  
st_crs(climatologies_europe_eobs_1950_2022_lon_lat_sf) <- 4326

# Find out closest coordinates indexes
climatologies_europe_eobs_1950_2022_lon_lat_nearest_coords <- st_nearest_feature(cities_set_sf, climatologies_europe_eobs_1950_2022_lon_lat_sf)

# Get the values and reverting the longitudes to the original format. Because we will use them to filter the # whole original file
climatologies_europe_eobs_1950_2022_lon_lat_values <- climatologies_europe_eobs_1950_2022_lon_lat_sf  |>  
  slice(climatologies_europe_eobs_1950_2022_lon_lat_nearest_coords) |>
  select(longitude,latitude) |> 
  (\(x) list( lon = pull(x,longitude), lat = pull(x,latitude)))()
  

rr <- climatologies_europe_eobs_1950_2022_lon_lat_sf  |>  
  slice(climatologies_europe_eobs_1950_2022_lon_lat_nearest_coords) |>
  select(longitude,latitude) 

# Add closest coords in original format to our set of cities

cities_set_ex_df <- add_column(cities_set_ex_df, lon_orig_eobs = climatologies_europe_eobs_1950_2022_lon_lat_values$lon, lat_orig_eobs = climatologies_europe_eobs_1950_2022_lon_lat_values$lat)

get_time_slices_eobs_for_cities <- function(cities_set_ex_df, climatologies_eobs_nc) {
#' Get time slices from the gridded climatologies netCDF file for the cties in cities_set
#' 
#' Return an sf with simple points for all the coordinates, and some merged info from the cities set  
#'

  climatology_cities_dat <- cities_set_ex_df$city |> 
    map(function(city_name) {
          lon = cities_set_ex_df |> filter(city == city_name) |> pull(lon_orig_eobs) 
          lat = cities_set_ex_df |> filter(city == city_name) |> pull(lat_orig_eobs)
          print(city_name)
          climatologies_eobs_nc |> 
            hyper_filter( longitude =  longitude == lon, latitude = latitude == lat) |>
            hyper_tibble() |>
            add_column(city = city_name)
        }) |>
    bind_rows() |>
    mutate(t2m_c = tg) |> 
    mutate(date_time = as_datetime(c(time * 60 * 60 * 24), origin="1950-01-01")) |>
    mutate(year = year(date_time)) |>
    mutate(month = month(date_time)) |>
    mutate(month_name = month(date_time, label = TRUE)) |>
    mutate(decade = as.character(as.integer((year %/% 10) * 10))) |> # Add decade class
    st_as_sf(coords = c("longitude", "latitude"),remove = FALSE)

  st_crs(climatology_cities_dat) <- 4326
  climatology_cities_dat
}

# Convert the anomalies to absolute temperature values
climatologies_europe_eobs_1950_2022_dat <- 
  get_time_slices_eobs_for_cities(cities_set_ex_df, climatologies_europe_eobs_1950_2022_nc) 

hh <- climatologies_europe_eobs_1950_2022_dat |>
  filter(city = "Santa Cruz")

# Combine all ERA5 dats
climatology_cities_dat <- bind_rows(climatology_cities_pre_2022_dat, climatology_cities_2022_dat)

# Calculate BaseLine Mean Annual Temperature for each city (1990-2020) 
tt <- climatologies_europe_eobs_1950_2022_dat |> 
  filter(year >= 1981 & year <= 2010) |>
  group_by(city) |>
  summarise(temperature_annual_mean = mean(t2m_c))
   
# Add Preindustrial baseline values for city, month
climatology_cities_dat <- climatology_cities_dat |>
  left_join(pre_industrial_baseline_dat, by = c("city", "month"))

climatology_cities_t2m_c <- climatology_cities_dat |> 
  group_by(city, month) %>%
  summarise(t2m_c_mean = mean(t2m_c))

climatology_cities_1990_2020_t2m_c <- climatology_cities_dat |>
  filter(year >= 1990 & year <= 2020) |>
  group_by(city, month) %>%
  summarise(t2m_c_mean = mean(t2m_c))

climatology_cities_decade_t2m_c <- climatology_cities_dat |>
  group_by(city, month, month_name, decade) %>%
  summarise(t2m_c_mean = mean(t2m_c))
  
  
```


## Plotting Data

Showing the time series of the max temperature at the 17:00 hour of the day

```{r echo=FALSE}
ggplot(climatology_cities_t2m_c, aes(month, t2m_c_mean, colour = city)) + 
  geom_line(stat = "identity") + 
  facet_wrap(~city, ncol = 1)

ggplot(climatology_cities_1990_2020_t2m_c, aes(month, t2m_c_mean, colour = city)) + 
  geom_line()


climatology_cities_decade_t2m_c |>
  filter(decade == "195" | decade == "202") |>
  ggplot(aes(month, t2m_c_mean, colour = decade)) + 
  geom_line() +
  facet_wrap(~city)

climatology_cities_decade_t2m_c |>
  filter(city == "Madrid", decade == "195" | decade == "196" | decade == "197" |  decade == "201" | decade == "202") |>
  ggplot(aes(month, t2m_c_mean, colour = decade)) + 
  geom_line(stat = "identity") +
  scale_x_date(, date_labels = "%b-%y")


climatology_cities_dat |>
  filter(city == "Madrid", month == 7) |>
  ggplot(aes(year, t2m_c, colour = decade)) + 
  geom_line(stat = "identity")

```

Progress of temperature in July in Madrid

```{r echo=FALSE}

plot_ele <- climatology_cities_dat |>
  filter(city == "Madrid", month == 7) |>
  ggplot(aes(year, t2m_c, colour = decade)) + 
  theme(plot.title = element_text(size = 14))

plot_ele +
  ggtitle("July in Madrid - The frog Experiment") +
  geom_line(stat = "identity") +
  geom_point() +
 labs(
    x = "Year", 
    y = "Average temperature Kº", 
    colour = "Decade",
    title = "July in Madrid - The frog Experiment",
    subtitle = "Source: Contains modified Copernicus Climate Change Service information from ERA5 monthly averaged data on single levels from 1959 to present[1959-2022]"
  )
```
Tale of two Cities in July

```{r}

# Create out directories
graph_out_dir <- "./out" 

graph_main_dir <- "/story_of_two_cities_july"

graph_main_dir_path <- path(graph_out_dir, graph_main_dir)

graphs_svg_sub_dir_path = path(graph_main_dir_path, "/svg")

graphs_plotly_sub_dir_path = path(graph_main_dir_path, "/plotly")

dir_create(path(graphs_svg_sub_dir_path))
dir_create(path(graphs_plotly_sub_dir_path))

graphs_path_files = c()
          
#
cities_in_study = c("Sevilla", "Burgos")
month_in_study = 7.
year_in_middle = median(climatology_cities_dat$year)

# Prepare Name of columns (for tooltips)
climatology_cities_ext_dat <- climatology_cities_dat |>
  mutate(City = city, Year = year, "Temp Avgr." = t2m_c )

labels_dat <- cities_in_study |> map(function(city_name) {
    climatology_cities_ext_dat |> 
    filter(city == city_name, year == year_in_middle, month == month_in_study) |>
    pull(t2m_c)
  }) |> 
  unlist() |>
  (\(y_values) {
    print(y_values)
   tibble( city = cities_in_study, x = year_in_middle, y = y_values)
})()


story_of_two_cities_july_base_plot <- climatology_cities_ext_dat |>
  filter(city %in% cities_in_study, month == 7) |>
  ggplot(aes(x = Year, y = `Temp Avgr.`, group = City)) + 
  theme(plot.title = element_text(size = 14)) +
   labs(
    x = "Year", 
    y = "Average temperature Cº", 
    title = "Average Temperature in July - Tale of two Cities? or the same story..",
    caption = "Data from: CDS - ERA5 monthly averaged data on single levels from 1959 to present[1959-2022]"
  )

story_of_two_cities_july_plot = story_of_two_cities_july_base_plot +
  geom_point() +
  geom_line(stat = "identity") +
  geom_smooth(method = "lm", se = FALSE)   +
  geom_label(data = labels_dat, alpha = 0.75, inherit.aes = FALSE, aes(x, y, label = city))

story_of_two_cities_july_plot

# Save to svg
svg_file_path <- path(graphs_svg_sub_dir_path, "/story_of_two_cities_july_plot", ext = "svg")
svglite(svg_file_path)
story_of_two_cities_july_plot
invisible(
      dev.off()
)

graphs_path_files <- c(graphs_path_files, svg_file_path)

# Same graph interactive Plotly
story_of_two_cities_july_plotly <- ggplotly(story_of_two_cities_july_plot) |> 
  add_annotations(
    x = labels_dat$x,
    y = labels_dat$y,
    text = labels_dat$city,
    xref = "x",
    yref = "y",
    showarrow = TRUE,
    arrowhead = 1,
    arrowsize = 0.5,
    arrowside = "none",
    ax = 20,
    ay = -40,
    bgcolor="white",
    opacity=0.8
  )

plotly_file_path <- path(graphs_svg_sub_dir_path, "/story_of_two_cities_july_plot", ext = "html")
htmlwidgets::saveWidget(
  story_of_two_cities_july_plotly, 
  plotly_file_path
)

story_of_two_cities_july_plotly

graphs_path_files <- c(graphs_path_files, plotly_file_path)


#########################################
#
# Genneate small html with links
#
#########################################


html_graph_links <- graphs_path_files |> map(function(file_path) {
  str_interp('<a href="${path(graph_main_dir, path_file(file_path))}"> A graph</a>')
  }) |> unlist() |> paste(collapse='\n')


html_page = paste(
  '
<!DOCTYPE html>
<html lang="en">
  <head>
    <meta charset="UTF-8">
    <meta name="viewport" content="width=device-width, initial-scale=1.0">
    <meta http-equiv="X-UA-Compatible" content="ie=edge">
    <title>Graphs</title>
  </head>
  <body>',
  html_graph_links,
'  </body>
' , 
  sep = "\n"
)

write_file(html_page, path(graph_out_dir, "/index", ext = "html"))


```
Tale of two Cities - Whole Year

```{r echo=FALSE}

cities_in_study = c( "Burgos", "Oviedo")


labels_dat <- climatology_cities_dat |>
  filter(city %in% cities_in_study, year == 1990) 
  
plot_ele <- climatology_cities_dat |>
  filter(city %in% cities_in_study, year == 1980) |>
  ggplot(aes(date_time, t2m_c, group = city,  colour = city)) + 
  theme(plot.title = element_text(size = 14)) +
   labs(
    x = "Year-Month", 
    y = "Average temperature Kº", 
    title = "(July) Tale of two Cities? or the same story..",
    subtitle = "Source: Contains modified Copernicus Climate Change Service information from ERA5 monthly averaged data on single levels from 1959 to present[1959-2022]"
  )

story_of_two_cities_july_plot <- plot_ele +
  geom_line(stat = "identity") +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
  # geom_label(data = labels_dat, aes(label = city)) 

story_of_two_cities_july_plot
```





Tale of two Cities in July II

```{r echo=FALSE}

climatology_cities_ext_dat <- climatology_cities_dat |>
  filter(city == "Sevilla" | city == "Gijón", month == 7) |>
  group_by(city, decade) |>
  mutate(temperature_decade_mean_c = mean(t2m_c)) 


labels_dat <- climatology_cities_dat |>
  filter(city == "Sevilla" | city == "Gijón", month == 7, year == 1990) 
  
plot_ele <- climatology_cities_ext_dat |>
  filter(city == "Sevilla" | city == "Gijón", month == 7) |>
  ggplot(aes(year, t2m_c, colour = decade, group = city)) + 
  theme(plot.title = element_text(size = 14)) +
   labs(
    x = "Year", 
    y = "Average temperature Cº", 
    colour = "Decade",
    title = "(July) Tale of two Cities? or the same story..",
    subtitle = "Source: Contains modified Copernicus Climate Change Service information from ERA5 monthly averaged data on single levels from 1959 to present[1959-2022]"
  )

story_of_two_cities_july_plot <- plot_ele +
  geom_line(stat = "identity") +
  geom_point() +
  # geom_smooth(method = "lm") +
  geom_label(data = labels_dat, aes(label = city)) +
  geom_line(aes(year, temperature_decade_mean_c, colour = decade, group=interaction(city, decade)))

story_of_two_cities_july_plot
```




Saving Plot as SVG

```{r echo=FALSE}

svglite("./out/story_of_two_cities_july_plot.svg")
story_of_two_cities_july_plot
invisible(
      dev.off()
)

```






Progress of temperature in August in Madrid

```{r echo=FALSE}

climatology_cities_dat |>
  filter(city == "Madrid", month == 8) |>
  ggplot(aes(year, t2m_c, colour = decade)) + 
  geom_line(stat = "identity")

climatology_cities_dat |>
  filter(city == "Madrid", month == 8) |>
  ggplot(aes(year, t2m_c, colour = decade)) + 
  geom_point()

```


```{r echo=FALSE}
p <- ggplot(mtcars, aes(wt, disp)) + geom_point()
p + facet_wrap(vars(vs,am))
```


Interactive


```{r echo=FALSE}

points.df <- data.frame(
  'x' = c(-53.50000, -54.15489, -54.48560, -52.00000, -52.57810, -49.22097, -48.00000),
  'y' = c(-38.54859, -39.00000, -38.80000, -38.49485, -38.00000, -38.50000, -37.74859)
  )


line.df <- data.frame(
  'x' = c(-52.53557, -52.00000, -50.00000, -48.00000, -46.40190),
  'y' = c(-41.00000, -40.60742, -40.08149, -40.82503, -39.00000)
  )

# Create 'sf' objects -----------------------------------------------------

points.sf <- st_as_sf(points.df, coords = c("x", "y"))

st_crs(points.sf) <- st_crs(4326) # assign crs
points.sf <- st_transform(points.sf, crs = 32721) # transform

line.sf <- st_sf(id = 'L1', st_sfc(st_linestring(as.matrix(line.df), dim = "XY")))
st_crs(line.sf) <- st_crs(4326) # assign crs
line.sf <- st_transform(line.sf, crs = 32721) # transform


# Plots -------------------------------------------------------------------

xmin <- min(st_bbox(points.sf)[1], st_bbox(line.sf)[1])
ymin <- min(st_bbox(points.sf)[2], st_bbox(line.sf)[2])
xmax <- max(st_bbox(points.sf)[3], st_bbox(line.sf)[3])
ymax <- max(st_bbox(points.sf)[4], st_bbox(line.sf)[4])

plot(points.sf, pch = 19, col = "#53A8BD", xlab = "Longitude", ylab = "Latitude",
     xlim = c(xmin,xmax), ylim = c(ymin,ymax), graticule = st_crs(4326), axes = TRUE)
plot(line.sf, col = "#C72259", add = TRUE)
text(st_coordinates(points.sf), as.character(1:7), pos = 3)
text(st_coordinates(line.sf), LETTERS[1:5], pos = 1)


distances <- st_distance(points.sf, line.sf)
print(distances)

```



