---
title: "Average Temperature and Precipitation on main cities"
output: html_document
date: '2023-09-24'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup 

ECMWF services

Follow the instructions in https://bluegreen-labs.github.io/ecmwfr/  in the section "Use: ECMWF services" to register and setup the key in your key-chain.
Also you need to agree on some terms of the institution following a link in the instructions of the above webpage.

CDS service:

Follow the instructions on the link in the section "Use: Copernicus Climate Data Store (CDS)".

Also you need to agree on some terms of the institution following a link in the instructions of the above webpage.

## Libraries

```{r echo=FALSE}

library("ecmwfr")
library(sf)
library(ncdf4)
library(tidyverse)
library(lubridate, warn.conflicts = FALSE)
library(tidync)

library(ggplot2)
library(hrbrthemes)

library(plotly)

```

## Getting Data

We have generated the data from the CDS web site and store the results on the file -> 

```{r echo=FALSE}

# 'ERA-5' reanalysis data for Madrid City Area

request <- list(
  product_type = "reanalysis",
  variable = c(
    "2m_temperature",
    "maximum_2m_temperature_since_previous_post_processing",
    "minimum_2m_temperature_since_previous_post_processing",
    "sea_surface_temperature"
  ),
  year = "1959",
  month = c(
    "01",
    "02",
    "03",
    "04",
    "05",
    "06",
    "07",
    "08",
    "09",
    "10",
    "11",
    "12"
  ),
  day = c(
    "01",
    "02",
    "03",
    "04",
    "05",
    "06",
    "07",
    "08",
    "09",
    "10",
    "11",
    "12",
    "13",
    "14",
    "15",
    "16",
    "17",
    "18",
    "19",
    "20",
    "21",
    "22",
    "23",
    "24",
    "25",
    "26",
    "27",
    "28",
    "29",
    "30",
    "31"
  ),
  time = c(
    "17:00"
  ),
  area = c(40.52,-3.82, 40.3,-3.53), # Aproximate Madrid City Area
  format = "netcdf",
  dataset_short_name = "reanalysis-era5-single-levels",
  target = "madrid-area-temperature_1959-1700.nc"
)

# If you have stored your user login information
# in the keyring by calling cds_set_key you can
# call:
file <- wf_request(
 user     = "150733",   # user ID (for authentification)
 request  = request,  # the request
 transfer = TRUE,     # download the file
 path     = "."       # store data in current working directory
 )

```

## Loading data

Reading data from ncdf file

```{r echo=FALSE}

# Create data structure

mad_temperature_nc <- tidync("./madrid-area-temperature_1959.nc")

mad_temperature_init_tbbl <- mad_temperature_nc %>% hyper_tibble()

mad_temperature_tbbl <- 
  mad_temperature_init_tbbl %>%
  mutate(time_date_var = as_datetime(c(time * 60 * 60), origin="1900-01-01"))

# Mean temperature values over the few grid points of Madrid Area
mad_temperature_mean_tbbl <- 
  mad_temperature_tbbl %>%
  group_by(time, time_date_var) %>%
  summarise(t2m_mean = mean(t2m), sst_mean = mean(sst), mx2t_mean = mean(mx2t), mn2t_mean = mean(mn2t), n = n())

```


## Plotting Data

Showing the time series of the max temperature at the 17:00 hour of the day

```{r echo=FALSE}

p <- ggplot(mad_temperature_mean_tbbl, aes(x=time_date_var, y=mx2t_mean)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("") +
  theme_ipsum() +
  theme(axis.text.x=element_text(angle=60, hjust=1))
  # scale_x_date(limit=c(as.Date("2017-01-01"),as.Date("2017-02-11"))) +
  # ylim(0,1.5)

p



scatterPlot <- 
  mad_temperature_mean_tbbl %>% 
  ggplot(aes(x=time_date_var, y=mx2t_mean)) + 
  geom_point(alpha=0.7, colour = "#51A0D5") + 
  labs(x = "Date Time", 
       y = "Max Temperature at 17:00",
       title = "Temperature in Madrid City Area") +
  theme_classic()

ggplotly(scatterPlot)
```
Interactive


```{r echo=FALSE}

# Turn it interactive with ggplotly
p <- ggplotly(p)
p

```



